---
title: How to make beautiful county maps of the US
date: '2019-10-07'
slug: how-to-make-county-maps
categories:
  - resources
tags:
  - R
  - data
  - code
  - tidycensus
subtitle: 'Using the R package Tidycensus'
summary: ''
lastmod: '2019-10-06T18:43:58-06:00'
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
---

```{r, include=F}

knitr::opts_chunk$set(echo = FALSE, warning = F, message = F, error = F)

require(tidycensus)
require(tidyverse)
require(viridis)

census_api_key("ba969e2db82634bf9988d89ae120f4b0f2955188")

```

First, you need to obtain an API key — it's quick and easy. [Go here](https://api.census.gov/data/key_signup.html), fill in the short form, and they'll give you one immediately. You need to plug in your API key where you see XXXXXX below.

To find variables of interest, I recommend the [Census website's advanced search](https://factfinder.census.gov/faces/nav/jsf/pages/searchresults.xhtml?refresh=t).

```{r, echo=TRUE}

require(tidycensus)
require(tidyverse)
require(viridis)

# Plug in your API key where the X's are; then uncomment that line.

# census_api_key("XXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXXX")

#################################
## Using the tidycensus package 
#################################

us.county.income <- get_acs(geography = "county", variables = "B19013_001", 
                            shift_geo = TRUE, geometry = TRUE)

ggplot(us.county.income) + 
  geom_sf(aes(fill = estimate), color = NA) + 
  coord_sf(datum = NA) + 
  theme_minimal() + 
  scale_fill_viridis_c() +
  labs(fill = "Estimated Avg.") 

```

For more advanced instructions, including some tips on how to search the Census variables from within R, see [these slides by Kyle Walker.](https://factfinder.census.gov/faces/nav/jsf/pages/searchresults.xhtml?refresh=t)