---
title: Make interactive graphs of financial time-series in five minutes with R (Bitcoin example)
date: '2019-10-15'
slug: analyze-financial-time-series-without-touching-a-spreadsheet
categories:
  - resources
tags:
  - data
  - R
subtitle: ''
summary: ''
authors: []
featured: no
image:
  caption: ''
  focal_point: ''
  preview_only: no
projects: []
---


```{r, include=F}

knitr::opts_chunk$set(echo = TRUE, warning = F, message = F, error = F)

```

We'll use two packages: The package `Quandl` for programmatically accessing a variety of financial time-series, and the package `dygraphs` for lovely interactive graphs based on the eponymous Javascript library. If you don't already have them, install them.

Below, we load the packages and run a search through the Quandl database to see what kind of variables we have to choose from. Feel free to explore your own interests and run the following example on a different variable. For this example, I'm interested in Bitcoin so I'll use the `Quandl.search()` function to search for variables mentioning `bitcoin`. When you find a result you're interested in, note the dataset code and the database code.

```{r}

require(dygraphs)
require(Quandl)

search.results <- Quandl.search("bitcoin", silent=T)

head(search.results, 2) # Show the first 2 results

```


Next, we'll use the `Quandl()` function to import the time-series of interest. I'm selecting the `BTC24USD` dataset from the `BCHARTS` database, as revealed to me in the search results above. Set the `type` option to `xts,` which is just one of R's widely-used formats for time-series variables. It plays well with the dygraph package.

```{r}

bitcoin <- Quandl("BCHARTS/BTC24USD", type = "xts")

```

After running the above line, you should now have a dataframe object named `bitcoin.` To produce a nice, interactive plot, simply use the `dygraph()` function and specify the variable of interest from your dataframe. Give it a title, and run it.

```{r}

dygraph(bitcoin$`Weighted Price`, main = "Bitcoin Exchange Rate (BTC vs. USD")

```

<br>

Pretty slick, eh? And you didn't even *touch* a spreadsheet, let alone Javascript.

What would you like to do next? Let me know. Forecast future prices? Test a hypothesis about what drives the value up/down? Add additional time-series to one graph? Reply or email me.

Hat tip to [Kyle Walker](https://walkerke.github.io/2014/12/charting-time-series-data-with-dygraphs-in-r-and-python/).
